\documentclass{charun}
\title{Git notes, version 0.13.1}
\author{Leon Suwalski}

\usepackage{xcolor}
% copied from https://www.color-hex.com/color-palette/85828
\definecolor{gitcolor}{HTML}{ffecec}
\definecolor{gitcolor2}{HTML}{990000}

% copied from https://tex.stackexchange.com/questions/132849/how-can-i-change-the-font-size-of-the-number-in-minted-environment
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}[breakable,boxsep=5pt,left=0pt,right=0pt,top=0pt,bottom=0pt,boxrule=0.5pt,arc=0pt,outer arc=0pt,colback=gitcolor,colframe=gitcolor2]\small}%
\AfterEndEnvironment{minted}{\end{tcolorbox}}%
    
\newcommand{\git}[1]{\marginpar[\textbf{\color{blue}{#1}}]{\textbf{\color{blue}{#1}}}}

\usepackage{marginnote}

\begin{document}
\maketitle
\raggedright

\textbf{git} (v2.45.2) is a stupid content tracker.

\section{Git commands -- configuration}
\input{sections/git-config}

\section{Git commands -- setting up a repository}
\input{sections/git-init}
\input{sections/git-clone}

\section{Git commands -- tracking changes}
\input{sections/git-add}
\input{sections/git-commit}
\input{sections/git-stash}
\input{sections/git-mv}
\input{sections/git-rm}

\section{Git commands -- branching}
\input{sections/git-branch}
\input{sections/git-checkout}
\input{sections/git-merge}
\input{sections/git-merge-base}
\input{sections/git-rebase}
% TODO: reflog
\input{sections/git-cherry-pick}

\section{Git commands -- tags}
\input{sections/git-tag}

\section{Git commands -- history}
\input{sections/git-status}
\input{sections/git-log}
\input{sections/git-shortlog}
\input{sections/git-show}
\input{sections/git-rev-parse}
\input{sections/git-blame}
\input{sections/git-diff}

\section{Git commands -- undoing changes}
\input{sections/git-revert}
\input{sections/git-clean}
\input{sections/git-reset}
\input{sections/git-gc}

\section{Git commands -- collaborating}
\input{sections/git-remote}
\input{sections/git-push}
\input{sections/git-pull}
\input{sections/git-fetch}

\section{Specifying revisions}
\begin{minted}{bash}
git show HEAD^  # (first) parent commit
git show HEAD^^ # second parent, useful for merge commits
git show HEAD^2 # second parent
git show HEAD~  # (first) parent commit
git show HEAD~~ # first parent of first parent
git show HEAD~2 # first parent of first parent

git log master..experiment
# assuming experiment branched out from master,
# this lists experiment commits that are not on master

git log ^master experiment
git log experiment --not master
# same as above

git log master...experiment
# this lists commits present on one but not both branches

git log --left-right master...experiment
# shows < before master and > before experiment commits


# @{u} refers to the upstream branch which the
# current branch is tracking, for example:
git merge @{u}
# @{upstream} means the same.
\end{minted}

\section{To be done}
To remove a commit:
\begin{minted}{bash}
git rebase --onto ~<sha>
\end{minted}

To remove a submodule:
\begin{minted}{bash}
git submodule deinit <name>
git rm <name>
git rm --cached <name>
rm -rf .git/modules/<name>
\end{minted}

To remove a tag:
\begin{minted}{bash}
git tag -d <tag>
git push <remote> :refs/tags/<tag>
\end{minted}

To undo a change:
\begin{minted}{bash}
git reset HEAD -- # unstage
git reset --mixed HEAD~ # uncommit
git checkout -- # discard
git restore
\end{minted}

% git checkout . undoes unstaged local modification
% git reset undoes staged modifications

\texttt{fileMode false} don't honor the executable bit of files in the working tree.

\section{.gitignore (work in progress)}
\begin{enumerate}
\item A blank line matches no files, so it can serve as a separator for readability.
\item A line starting with \# serves as a comment. Put a backslash ("\") in front of the first hash for patterns that begin with a hash.
\item Trailing spaces are ignored unless they are quoted with backslash ("\").
\item An optional prefix "!" which negates the pattern; any matching file excluded by a previous pattern will become included again. It is not possible to re-include a file if a parent directory of that file is excluded. Git doesn’t list excluded directories for performance reasons, so any patterns on contained files have no effect, no matter where they are defined. Put a backslash ("\") in front of the first "!" for patterns that begin with a literal "!", for example, "\!important!.txt".
\item The slash "/" is used as the directory separator. Separators may occur at the beginning, middle or end of the .gitignore search pattern.
\item If there is a separator at the beginning or middle (or both) of the pattern, then the pattern is relative to the directory level of the particular .gitignore file itself. Otherwise the pattern may also match at any level below the .gitignore level.
\item If there is a separator at the end of the pattern then the pattern will only match directories, otherwise the pattern can match both files and directories.
\item For example, a pattern doc/frotz/ matches doc/frotz directory, but not a/doc/frotz directory; however frotz/ matches frotz and a/frotz that is a directory (all paths are relative from the .gitignore file).
\item An asterisk "*" matches anything except a slash. The character "?" matches any one character except "/". The range notation, e.g. [a-zA-Z], can be used to match one of the characters in a range. See fnmatch(3) and the \mintinline{bash}{FNM_PATHNAME} flag for a more detailed description.
\item Two consecutive asterisks ("**") in patterns matched against full pathname may have special meaning:
\item A leading "**" followed by a slash means match in all directories. For example, "**/foo" matches file or directory "foo" anywhere, the same as pattern "foo". "**/foo/bar" matches file or directory "bar" anywhere that is directly under directory "foo".
\item A trailing "/**" matches everything inside. For example, "abc/**" matches all files inside directory "abc", relative to the location of the .gitignore file, with infinite depth.
\item A slash followed by two consecutive asterisks then a slash matches zero or more directories. For example, "a/**/b" matches "a/b", "a/x/b", "a/x/y/b" and so on.
\item Other consecutive asterisks are considered regular asterisks and will match according to the previous rules.
\end{enumerate}

Todo: git ls-remote, GPG,

\section{To be done later}
Distributed git: git request-pull, git format-patch -M, git send-email.
git diff \$(git merge-base contrib master) equals git diff master...contrib.
% „gpg -a --export F721C45A | git hash-object -w --stdin”
git describe.


\end{document}

Done:
https://stackoverflow.com/questions/tagged/git

To do:
https://stackoverflow.com/questions/tagged/git?tab=votes&page=2&pagesize=15