%

\textbf{git log}\marginpar[\textbf{\color{blue}{log}}]{\textbf{\color{blue}{log}}} displays entire commit history (without any diff output by default).

\mintinline{bash}{git log foo..bar} displays only commits between \texttt{foo} and \texttt{bar}.
Similarly \mintinline{bash}{git log foo ^bar} displays only commits reachable from foo but not bar.

\mintinline{bash}{git log --author=<pattern> --committer=<pattern>} displays only commits with author/committer matching regex.

\mintinline{bash}{git log --grep=<pattern>} displays only commits with log message matching regex.

\mintinline{bash}{git log --follow <path>} continues listing the history beyond renames but works only for a single file!

\mintinline{bash}{git log --no-merges} skips commits with more than one parent.

\mintinline{bash}{git log --decorate} prints nice reference names without refs/heads/, refs/tags/ and refs/remotes/ prefixes.

\mintinline{bash}{git log --oneline} skips author, date and full commit message. Equivalent to \mintinline{bash}{--pretty=oneline --abbrev-commit}.

\mintinline{bash}{git log --graph} draws a GuitarHero-like text based graph.

\mintinline{bash}{git log --stat} generates a diffstat (histogram).

\mintinline{bash}{git log --format='format:%h %an'} uses custom format.
When using format option, following placeholders are commonly used:
\mintinline{bash}{%H} (commit hash),
\mintinline{bash}{%h} (abbreviated commit hash),
\mintinline{bash}{%an} or \mintinline{bash}{%aN} (author name),
\mintinline{bash}{%ae} or \mintinline{bash}{%aE} (author e-mail),
\mintinline{bash}{%ad} or \mintinline{bash}{%aD} or \mintinline{bash}{%ar} or \mintinline{bash}{%at} or \mintinline{bash}{%ai} or \mintinline{bash}{%aI} or \mintinline{bash}{%as} or \mintinline{bash}{%ah} (author date),
\mintinline{bash}{%s} and \mintinline{bash}{%b} (subject and body).
Author placeholders have corresponding committer placeholders.

Various date formats:
\begin{compactenum}
\item [\texttt{\%ad}] $\to\,\,$ \texttt{Sat Sep 18 19:40:37 2021}
\item [\texttt{\%aD}] $\to\,\,$ \texttt{Sat, 18 Sep 2021 13:40:37 -0400}
\item [\texttt{\%ah}] $\to\,\,$ \texttt{Sep 18 2021}
\item [\texttt{\%ai}] $\to\,\,$ \texttt{2021-09-18 13:40:37 -0400}
\item [\texttt{\%aI}] $\to\,\,$ \texttt{2021-09-18T13:40:37-04:00}
\item [\texttt{\%ar}] $\to\,\,$ \texttt{1 year, 8 months ago}
\item [\texttt{\%as}] $\to\,\,$ \texttt{2021-09-18}
\item [\texttt{\%at}] $\to\,\,$ \texttt{1631986837}
\end{compactenum}

Useful short options:
\begin{compactenum}
\item [\texttt{-G}] looks for diffs whose patch contains lines matching regex.
\item [\texttt{-n}] limits the number of commits to output (equivalent to \mintinline{bash}{--max-count=}),
\item [\texttt{-p}] generates a patch (equivalent to \mintinline{bash}{--patch}),
\item [\texttt{-S}] looks for diffs that change \# of occurrences of given string.
\item [\texttt{-U}] generates diffs with ... instead of 3 lines of context (equivalent to \mintinline{bash}{--unified=})
\end{compactenum}

%