\subsection{klasa String}
% https://ruby-doc.org/3.3.6/String.html#method-i-split
Obiekt tej klasy jest ciągiem znaków, zazwyczaj przedstawiającym tekst.
Niektóre metody bez wykrzykników mutują argumenty, na przykład \mintinline{ruby}{String#replace}.
Klasa \mintinline{ruby}{String} zawiera moduł \mintinline{ruby}{Comparable}.

\subsubsection{Methods for Querying}
\textbf{length} (znane też jako \textbf{size}) liczy znaki, nie bajty (porównaj z klasą \mintinline{ruby}{Enumerable}).

empty?: Returns true if self.length is zero; false otherwise.

bytesize: Returns the count of bytes.

count: Returns the count of substrings matching given strings.

Substrings

% #=~: Returns the index of the first substring that matches a given Regexp or other object; returns nil if no match is found.

index: Returns the index of the first occurrence of a given substring; returns nil if none found.

rindex: Returns the index of the last occurrence of a given substring; returns nil if none found.

include?: Returns true if the string contains a given substring; false otherwise.

match: Returns a MatchData object if the string matches a given Regexp; nil otherwise.

match?: Returns true if the string matches a given Regexp; false otherwise.

\textbf{start\_with?} (\textbf{end\_with?}) zwraca, czy jeden łańcuch znaków zaczyna się (kończy się) pewnym innym łańcuchem.
\begin{minted}{ruby}
'hello'.start_with?('hell') # => true
'hello'.start_with?(/H/i) # => true
'hello'.start_with?('heaven', 'hell') # => true
'hello'.start_with?('heaven', 'paradise')
'hello'.end_with?('ello') # => true
'hello'.end_with?('heaven', 'ello') # => true
'hello'.end_with?('heaven', 'paradise')
\end{minted}

Encodings

encoding: Returns the Encoding object that represents the encoding of the string.

unicode\_normalized?: Returns true if the string is in Unicode normalized form; false otherwise.

valid\_encoding?: Returns true if the string contains only characters that are valid for its encoding.

ascii\_only?: Returns true if the string has only ASCII characters; false otherwise.

Other

sum: Returns a basic checksum for the string: the sum of each byte.

hash: Returns the integer hash code.

Methods for Comparing
==, ===: Returns true if a given other string has the same content as self.

eql?: Returns true if the content is the same as the given other string.

% #<=>: Returns -1, 0, or 1 as a given other string is smaller than, equal to, or larger than self.

casecmp: Ignoring case, returns -1, 0, or 1 as a given other string is smaller than, equal to, or larger than self.

casecmp?: Returns true if the string is equal to a given string after Unicode case folding; false otherwise.

Methods for Modifying a String
Each of these methods modifies self.

Insertion

insert: Returns self with a given string inserted at a given offset.

<<: Returns self concatenated with a given string or integer.

Substitution

sub!: Replaces the first substring that matches a given pattern with a given replacement string; returns self if any changes, nil otherwise.

gsub!: Replaces each substring that matches a given pattern with a given replacement string; returns self if any changes, nil otherwise.

succ!, next!: Returns self modified to become its own successor.

replace: Returns self with its entire content replaced by a given string.

reverse!: Returns self with its characters in reverse order.

setbyte: Sets the byte at a given integer offset to a given value; returns the argument.

tr!: Replaces specified characters in self with specified replacement characters; returns self if any changes, nil otherwise.

tr\_s!: Replaces specified characters in self with specified replacement characters, removing duplicates from the substrings that were modified; returns self if any changes, nil otherwise.


Encoding

encode!: Returns self with all characters transcoded from one given encoding into another.

unicode\_normalize!: Unicode-normalizes self; returns self.

scrub!: Replaces each invalid byte with a given character; returns self.

force\_encoding: Changes the encoding to a given encoding; returns self.

Deletion

clear: Removes all content, so that self is empty; returns self.

slice!, []=: Removes a substring determined by a given index, start/length, range, regexp, or substring.

squeeze!: Removes contiguous duplicate characters; returns self.

delete!: Removes characters as determined by the intersection of substring arguments.

lstrip!: Removes leading whitespace; returns self if any changes, nil otherwise.

rstrip!: Removes trailing whitespace; returns self if any changes, nil otherwise.

strip!: Removes leading and trailing whitespace; returns self if any changes, nil otherwise.

chomp!: Removes trailing record separator, if found; returns self if any changes, nil otherwise.

chop!: Removes trailing newline characters if found; otherwise removes the last character; returns self if any changes, nil otherwise.



\subsubsection{Methods for converting to new String}
Each of these methods returns a new String based on self, often just a modified copy of self.

Extension

*: Returns the concatenation of multiple copies of self,

+: Returns the concatenation of self and a given other string.

center: Returns a copy of self centered between pad substring.

concat: Returns the concatenation of self with given other strings.

prepend: Returns the concatenation of a given other string with self.

ljust: Returns a copy of self of a given length, right-padded with a given other string.

rjust: Returns a copy of self of a given length, left-padded with a given other string.

Encoding

b: Returns a copy of self with ASCII-8BIT encoding.

scrub: Returns a copy of self with each invalid byte replaced with a given character.

unicode\_normalize: Returns a copy of self with each character Unicode-normalized.


Substitution

dump: Returns a copy of self with all non-printing characters replaced by xHH notation and all special characters escaped.

undump: Returns a copy of self with all \\xNN notation replace by \\uNNNN notation and all escaped characters unescaped.

sub: Returns a copy of self with the first substring matching a given pattern replaced with a given replacement string;.

gsub: Returns a copy of self with each substring that matches a given pattern replaced with a given replacement string.

succ, next: Returns the string that is the successor to self.

\textbf{reverse} zwraca wszystkie znaki od tyłu (podobnie jak dla tablic).
Ma wersję z wykrzyknikiem.
\begin{minted}{ruby}
'stressed'.reverse # => "desserts"
\end{minted}

tr: Returns a copy of self with specified characters replaced with specified replacement characters.

tr\_s: Returns a copy of self with specified characters replaced with specified replacement characters, removing duplicates from the substrings that were modified.

%: Returns the string resulting from formatting a given object into self

\textbf{capitalize}, \textbf{downcase}, \textbf{upcase}, \textbf{swapcase} zmieniają wielkość liter.
Mają wersje z wykrzyknikiem.
\begin{minted}{ruby}
'hello World!'.capitalize # => "Hello world!"
'hello World!'.downcase   # => "hello world!"
'hello World!'.upcase     # => "HELLO WORLD!"
'hello World!'.swapcase   # => "HELLO wORLD!"
\end{minted}

Deletion

delete: Returns a copy of self with characters removed

delete\_prefix: Returns a copy of self with a given prefix removed.

delete\_suffix: Returns a copy of self with a given suffix removed.

lstrip: Returns a copy of self with leading whitespace removed.

rstrip: Returns a copy of self with trailing whitespace removed.

strip: Returns a copy of self with leading and trailing whitespace removed.

chomp: Returns a copy of self with a trailing record separator removed, if found.

chop: Returns a copy of self with trailing newline characters or the last character removed.

squeeze: Returns a copy of self with contiguous duplicate characters removed.

[], slice: Returns a substring determined by a given index, start/length, or range, or string.

byteslice: Returns a substring determined by a given index, start/length, or range.

chr: Returns the first character.

Duplication

to\_s, to\_str: If self is a subclass of String, returns self copied into a String; otherwise, returns self.




\subsubsection{Methods for converting to non-String}
Each of these methods converts the contents of self to a non-String.

Characters, Bytes, and Clusters

bytes: Returns an array of the bytes in self.

chars: Returns an array of the characters in self.

codepoints: Returns an array of the integer ordinals in self.

getbyte: Returns an integer byte as determined by a given index.

% grapheme_clusters: Returns an array of the grapheme clusters in self.

lines: Returns an array of the lines in self, as determined by a given record separator.

partition: Returns a 3-element array determined by the first substring that matches a given substring or regexp,

rpartition: Returns a 3-element array determined by the last substring that matches a given substring or regexp,

\textbf{split} returns an array of substrings determined by a given delimiter – regexp or string.
\begin{minted}{ruby}
'abc def ghi'.split(' ') # => ["abc", "def", "ghi"]
"abc \n\tdef\t\n  ghi".split(' ') # => ["abc", "def", "ghi"]
'abc  def   ghi'.split(' ') # => ["abc", "def", "ghi"]
''.split(' ') # => []
'abracadabra'.split('ab') # => ["", "racad", "ra"]
'aaabcdaaa'.split('a') # => ["", "", "", "bcd"]
''.split('a') # => []
'3.14159'.split('1') # => ["3.", "4", "59"]
'!@#$%^$&*($)_+'.split('$') # => ["!@#", "%^", "&*(", ")_+"]

# regexp delimeter
'abracadabra'.split(/ab/) # => ["", "racad", "ra"]
'aaabcdaaa'.split(/a/) # => ["", "", "", "bcd"]
'aaabcdaaa'.split(//) # => ["a", "a", "a", "b", "c", "d", "a", "a", "a"]
'1 + 1 == 2'.split(/\W+/) # => ["1", "1", "2"]

# with a limit - a positive integer
'aaabcdaaa'.split('a', 1) # => ["aaabcdaaa"]
'aaabcdaaa'.split('a', 2) # => ["", "aabcdaaa"]
'aaabcdaaa'.split('a', 5) # => ["", "", "", "bcd", "aa"]
'aaabcdaaa'.split('a', 7) # => ["", "", "", "bcd", "", "", ""]
'aaabcdaaa'.split('a', 8) # => ["", "", "", "bcd", "", "", ""]

# with a block
'abc def ghi'.split(' ') {|substring| p substring }
\end{minted}

Matching

scan: Returns an array of substrings matching a given regexp or string, or, if a block given, passes each matching substring to the block.

unpack: Returns an array of substrings extracted from self according to a given format.

unpack1: Returns the first substring extracted from self according to a given format.

Numerics

hex: Returns the integer value of the leading characters, interpreted as hexadecimal digits.

oct: Returns the integer value of the leading characters, interpreted as octal digits.

ord: Returns the integer ordinal of the first character in self.

% to_i: Returns the integer value of leading characters, interpreted as an integer.

% to_f: Returns the floating-point value of leading characters, interpreted as a floating-point number.

Strings and Symbols

inspect: Returns copy of self, enclosed in double-quotes, with special characters escaped.

% to_sym, intern: Returns the symbol corresponding to self.




Methods for Iterating
each\_byte: Calls the given block with each successive byte in self.

each\_char: Calls the given block with each successive character in self.

each\_codepoint: Calls the given block with each successive integer codepoint in self.

each\_grapheme\_cluster: Calls the given block with each successive grapheme cluster in self.

each\_line: Calls the given block with each successive line in self, as determined by a given record separator.

upto: Calls the given block with each string value returned by successive calls to succ.



