\section{Built-in Functions}
Logical operations:
\begin{minted}{python}
all() # False if any element is False
any() # False if all elements are False or iterable is empty
\end{minted}

\begin{itemize}
% \item \mintinline{python}{__import__()},

\item \mintinline{python}{abs()}: absolute value (of a number).

\item
\mintinline{python}{ascii()},

\item
\mintinline{python}{breakpoint()},
\mintinline{python}{bytearray()},
\mintinline{python}{bytes()},
\mintinline{python}{callable()},
\dotfill not implemented yet.

\item \mintinline{python}{chr()}: converts integer to character, for example \mintinline{python}{chr(97) == 'a'}.

\item \mintinline{python}{ord()}: converts character into integer, for example \mintinline{python}{ord('a') == 97}.

\item
\mintinline{python}{classmethod()},
\mintinline{python}{compile()},
\dotfill not implemented yet.
\end{itemize}

Converting:
\begin{minted}{python}
bool() # anything not listed below: True
# empty sequences, collections, zeroes, None, False: False

# hex, oct, bin require an integer
hex() # to a lowercase hexadecimal string prefixed with "0x"
oct() # to an octal string prefixed with "0o"
bin() # to a binary string prefixed with "0b"
       
# int, float require a number or string
int() # converts a number or string
float() # converts a number or string

complex() # converts a pair of numbers or string
# complex("3+4j") == complex(3,4) == (3+4j)

tuple()
list()
dict() # dictionary
frozenset()
set()
str() # string
\end{minted}

\begin{itemize}
\item
\mintinline{python}{delattr()},
\dotfill not implemented yet.

\item
\mintinline{python}{dir()},
\dotfill not implemented yet.

\item \mintinline{python}{divmod()}: quotient and remainder.

\item \mintinline{python}{enumerate()}: returns tuples containing a count (from 0) and consecutive values of a sequence or iterator.

\item
\mintinline{python}{eval()},
\mintinline{python}{exec()},
\dotfill not implemented yet.

\item \mintinline{python}{filter(function, iterable)}: returns only those elements for which function returns true.

\item \mintinline{python}{map(function, iterable)}: applies function to every item of iterable.

\item
\mintinline{python}{format()},
\mintinline{python}{getattr()},
\mintinline{python}{globals()},
\mintinline{python}{hasattr()},
\mintinline{python}{hash()},
\mintinline{python}{help()},
\mintinline{python}{id()},
\mintinline{python}{input()},
\mintinline{python}{isinstance()},
\mintinline{python}{issubclass()},
\mintinline{python}{iter()},
\dotfill not implemented yet.

\item \mintinline{python}{len()}: length (number of items) of an object.

\item
\mintinline{python}{locals()},
\dotfill not implemented yet.
\mintinline{python}{max()}: the largest item.
\mintinline{python}{min()}: the smallest item.

\item
\mintinline{python}{memoryview()},
\mintinline{python}{next()},
\mintinline{python}{object()},
\dotfill not implemented yet.

\item \mintinline{python}{open()}: opens a file and then closes matching file descriptor.
\begin{minted}{python}
with open("path/to/file") as f:
	for line in f:
		# do something

with open("path/to/file", "w") as f:
	f.write("some string")
\end{minted}

\item \mintinline{python}{pow()}: exponentiation, with optional parameter for modulus.

\item \mintinline{python}{print()}: :)

\item
\mintinline{python}{property()},
\dotfill not implemented yet.

\item \mintinline{python}{range()}: immutable sequence type...

\item
\mintinline{python}{repr()},
\dotfill not implemented yet.

\item \mintinline{python}{reversed()}: returns a reverse iterator.

\item \mintinline{python}{round()}: rounds to given precision (to integer by default).

\item
\mintinline{python}{setattr()},
\mintinline{python}{slice()},
\dotfill not implemented yet.

\item \mintinline{python}{sorted()}: returns a new sorted list.

\item
\mintinline{python}{staticmethod()},
\dotfill not implemented yet.

\item \mintinline{python}{sum()}: sums items of an iterable from left to right.
For some use cases, there are good alternatives: \mintinline{python}{"".join(sequence)} for strings; \mintinline{python}{math.fsum()} for floating numbers and \mintinline{python}{itertools.chain()} for a series of iterables.

\item
\mintinline{python}{super()},
\dotfill not implemented yet.

\item \mintinline{python}{type()}: returns the type of an object.

\item
\mintinline{python}{vars()},
\dotfill not implemented yet.

\item \mintinline{python}{zip()}: aggregates (into tuples) elements from each of the iterables.
\end{itemize}

%